### 2016-03-23 Draft.tex
Insert line break at the end of each sentence.  (By find-and-replacing all period-double-spaces with an added line break... may not be 100% effective).

Our model is formulated for single-visit removal-sampled point-count surveys at a set of sites $s=1,\dotso,S$ repeated across distinct sampling periods $r=1,\dotso,R$.  We present the model for exact time-to-detection data and also show how to modify it for interval-censored detection times.  For exact data, we use $n_{sr}^{(obs)}$ to denote the number of individuals observed during a survey, each with a time to detection $t_{srq}$, where $q = 1,\dotso,n_{sr}^{(obs)}$ indexes observed individuals.  For interval-censored data, we enumerate the intervals $i = 1,\dotso,I$ with the cutoff times for each interval written $C_i$, where $C_0 = 0 \text{ and } C_I = \text{the survey duration}$.  The number of animals counted during any interval is $n_{sri}$, the vector of counts for a survey is written $\vn_{sr}$, and $n_{sr}^{(obs)} = \sum_{i=1}^I n_{sri}$.

@@

We construct a hierarchical model of removal-sampled point-counts in three pieces: (a) a binomial data model, (b) a detection model based upon the cumulative distribution function (cdf) of a regression-modeled TTDD, and (c) an abundance model based upon Poisson regression.  Our model incorporates covariates and random effects for both abundance and detection via mixed effects linear regression.

\subsection{Data Model}
We assume that counts from different surveys are conditionally independent.  We also assume that the times to detection for all individuals at any given survey are independent, identically distributed according to a TTDD with cdf $F_T(t)$ and pdf $f_T(t)$.  The probability of detecting an animal is thus $p_{sr}^{(det)} = Pr(t_{srq} < C_I) = F_T(C_I)$ and the conditional distribution of observed detection times can be written $f_{T|det}(t|t_{srq}<C_I) = f_T(t) / F_T(C_I)$.  If the total number of animals present at the survey is $N_{sr}$, then the data model is:
\begin{align}\label{eq:datamodel}
n_{sr}^{(obs)} &\sim \mbox{Binomial}\left(N_{sr}, p_{sr}^{(det)}\right)\\
\label{eq:Tsrq}T_{srq} &\sim f_{T|det}(t|t_{srq}<C_I)
\end{align}
When detection times are censored, we replace Eq. \ref{eq:Tsrq} with $\vn_{sr} \sim \mbox{Multinomial}\left(n_{sr}^{(obs)}, \textbf{p}_{sr}\right)$, where $\textbf{p}_{sr}$ is the vector of interval-specific detection probabilities obtained by partitioning the conditional cdf $F_{T|det}(t|t_{srq}<C_I)$ at the interval cutoff times.  It follows that $\sum_{i=1}^I p_{sri} = 1$.

@@

% \adam{I have contemplated on and off including a discussion of broad priors and the limit behavior as $\varphi \to 0$, but I think it's not really on topic.  Note: now that I've seen van Dishoeck and Manntyniemi, perhaps I should worry about informativeness in priors.}

>>>>

Our model is formulated for single-visit removal-sampled point-count surveys at a set of sites $s=1,\dotso,S$ repeated across distinct sampling periods $r=1,\dotso,R$.  
We present the model for exact time-to-detection data and also show how to modify it for interval-censored detection times.  
We use $n_{sr}^{(obs)}$ to denote the number of individuals observed during a survey.  
For exact data, times to detection are $t_{srq}$, where $q = 1,\dotso,n_{sr}^{(obs)}$ indexes observed individuals.  
For interval-censored data, we enumerate the intervals $i = 1,\dotso,I$ with the cutoff times for each interval written $C_i$, where $C_0 = 0 \text{ and } C_I = \text{the survey duration}$.  
The number of animals counted during any interval is $n_{sri}$, the vector of counts for a survey is written $\vn_{sr}$, and $n_{sr}^{(obs)} = \sum_{i=1}^I n_{sri}$.

@@

We construct a hierarchical model of removal-sampled point-counts in three pieces: (a) a binomial data model, (b) a detection model based upon a TTDD, and (c) an abundance model based upon Poisson regression.  
Our model incorporates covariates and random effects for both abundance and detection via mixed effects linear regression.

\subsection{Data Model}
We assume that counts from different surveys are conditionally independent.  
We also assume that the times to detection for all individuals in any given survey are independent, identically distributed according to a TTDD with cdf $F_T(t)$ and pdf $f_T(t)$.  
The probability of detecting an animal is thus $p_{sr}^{(det)} = Pr(t_{srq} < C_I) = F_T(C_I)$ and the conditional distribution of observed detection times can be written $f_{T|det}(t|t_{srq}<C_I) = f_T(t) / F_T(C_I)$.  
Given $N_{sr}$ total number animals present at a survey, the data model is:
\begin{align}\label{eq:datamodel}
n_{sr}^{(obs)} &\sim \mbox{Binomial}\left(N_{sr}, p_{sr}^{(det)}\right)\\
\label{eq:Tsrq}T_{srq} &\sim f_{T|det}(t|t_{srq}<C_I)
\end{align}
When detection times are censored, we replace Eq. \ref{eq:Tsrq} with $\vn_{sr} \sim \mbox{Multinomial}\left(n_{sr}^{(obs)}, \textbf{p}_{sr}\right)$, where $\textbf{p}_{sr}$ is the vector of interval-specific detection probabilities obtained by partitioning the conditional cdf $F_{T|det}(t|t_{srq}<C_I)$ at the interval cutoff times.  
It follows that $\sum_{i=1}^I p_{sri} = 1$.

@@

% \adam{I have contemplated on and off including a discussion of broad priors and the limit behavior as $\varphi \to 0$, but I think it's not really on topic.  
% Note: now that I've seen van Dishoeck and Manntyniemi, perhaps I should worry about informativeness in priors.}
